---
title: "Bad plot and Good plot"
author: "Leah Barkai"
date: today
format: 
  html:
    theme: cosmo
    toc: true #table of contents
    toc-depth: 2
    code-fold: false #code is visible
    code-overflow: wrap #don't need to scroll in the code
editor: visual
knitr: #this is to save figures automatically 
  opts_chunk:
    fig.path: "../Output/" ## added the ../ because my qmd folder is in my scripts folder so need to go up one directory and then into my output folder
    message: false #gets rid of messages from code
    warning: false #gets rid of warnings from code
---

## Introduction

This document is for my homework for week 8's bad plot and good plot. I decided to use Leap Day dataset from a past Tidy Tuesdays. This dataset is available [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-02-27/readme.md). The dataset includes:

-   year_birth - Birth year of the person

-   person – name of the person

-   description - of who they are

-   year_death - year of their death

## Load Libraries

These are the libraries I needed for this homework assignment.

```{r}
library(tidyverse) #loads a bunch of data-science packages
library(knitr) #lets you run code chunks in Quarto and R Markdown
library(here) #to save my plot
library(ggimage) #To add in images to my plot
```

## Read in data

This data is from [Leap Day dataset](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-02-27/readme.md) from a past Tidy Tuesdays and the data came from the [February 29](https://en.wikipedia.org/wiki/February_29) article on Wikipedia.

> February 29 is a leap day (or "leap year day"), an intercalary date added periodically to create leap years in the Julian and Gregorian calendars.

```{r}
births <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-02-27/births.csv') #Load in the data - I decided to only do the births data not the deaths too
```

## Data Dictionary

#### `births.csv`

| variable    | class     | description                        |
|-------------|-----------|------------------------------------|
| year_birth  | integer   | Year in which this person was born |
| person      | character | Name of the person                 |
| description | character | Short description of the person    |
| year_death  | integer   | Year of death                      |

## Look at the data

This was to get an idea of what was in the data.

```{r}
glimpse(births) #look at the structure, types and first few rows of the data table
```

## Data Analysis

Need to make a summary table with number of births each year

```{r}
# Creating a summary table of number of births per year
births_count <- births %>% #new dataset name
  group_by(year_birth) %>% #grouping data by the variable - year_birth
  summarise(Number_of_births= n()) %>% #summary table with one row per year 
  arrange(year_birth) %>% #sorts table in ascending order by birth year
  rename(Birth_year= year_birth) %>% #Changing the column name to Birth_year
  select(Birth_year, Number_of_births) #Keep only these columns

births_count #View the summary table
```

### Bad Plot

This code below is to create my bad plot first

```{r bad_plot}
#| class.source: fill #Did this so I didn't have to scroll in the code block

#Adding new column to summary table for image and assign image to specific rows
births_count$image <- NA # new column with all NAs
births_count$image[c(5,10,15)] <- here("Good_plot_Bad_plot", "Data", "leaper.png") # assign image to rows 5, 10, 15
  #Image of person leaping is random, makes plot more cluttered and distracts from the data

#Bad plot code
bad_plot <- ggplot(births_count, #Making bad plot first
                   aes(x = Number_of_births, y = Birth_year)) + #this is what I want for the x and y axis 
    #Time should usually be on the x-axis
  
  geom_line(linewidth = 3, color = "red") + # thick red line inbetween data points
    #Makes it hard to see a trend or indidvual data points
  
  geom_point(aes(color = factor(Birth_year)), size = 5, alpha = 1) + #data points are colored by year 
  #makes the plot confusing and there is no legend to help you figure out which color equals what year
  
  geom_image(aes(image = image), size = 0.05) + #adds image to some points
    #only some points have the image, which makes it confusing
  
  theme(panel.background = element_rect(fill = "#FF6EC7"), #background color bright pink
        #Hard on the eyes
        panel.grid.major = element_line(color = "purple", linewidth = 2), #adds thick purple grid lines 
        #which makes it look kidish bc of the pink and purple and also hard to look at
        
  axis.text.x = element_text(angle = 90, #puts the x axis label at 90 degree angle 
                                   #Hard to read
                                   
                             hjust = 1, #right aligns text, putting to right of axis tick
                                   #making it more confusing
                                   
                             color = "green", #text color green 
                                   #making it hard to read on the pink background
                                   
                             size = 12), #large text size
                                   #making it distracting
        
  axis.text.y = element_text(color = "blue", size = 12)) + #y axis title blue and not the same as the x axis making it hard for the brain to look at
  
  labs(title = "LEAP DAY BIRTHS OVER TIME!!!", #Title of the graph 
       #all caps and many exclamation marks is too much 
       
       x = "Number of People", #title of the x-axis
       y = "Year of Birth", #title of the y-axis
       
       caption = "Data: TidyTuesday GitHub") + #Added caption to tell you where the data is from but the info given won't help you find it
  
  guides(color = "none") # removes the legend 
    #thus making more confusing 

#Saving my plot 
ggsave(here("Good_plot_Bad_plot","Output","Badplot.pdf"), plot = bad_plot, width = 8, height = 6) #saving my plot to the output folder

print(bad_plot) #viewing plot

```

### Good Plot

Now this code is for my good plot

```{r good_plot}
# Identify year with maximum number of births
max_births <- births_count %>% #new data frame
  filter(Number_of_births == max(Number_of_births)) # keep only rows with max births

# New data frame for the calendar image on where to put it
calendar_image <- data.frame( #new data frame
  x = min(births_count$Birth_year) + 5,  # X-coordinate is slightly right of left edge of plot
  y = 8.5, #y-coordinate is under title
  image = here("Good_plot_Bad_plot", "Data", "calendar_photo_leap_year.png") # calling the image
)

# Good plot
good_plot <- ggplot(births_count, aes(x = Birth_year, y = Number_of_births)) + #x-axis has the year bc it is a timeline plot
  
  geom_line(size = 1, color = "steelblue") +  #line in between data points
  geom_point(alpha = 0.8, color = "darkorange", size = 3) + #Data points color
  #Complementary colors and good for colorblind people
  
  geom_point(data = max_births, aes(x = Birth_year, y = Number_of_births), #Maximum birth point
             color = "red", size = 4) + #Put the dot in red to stand out
  geom_text(data = max_births, 
            aes(x = Birth_year, y = Number_of_births, label = Number_of_births), #Where the text goes 
            vjust = -0.5, #Puts it just above the point making it easy to read
            color = "red", fontface = "bold", size = 4) + #Making label stand out
  
  geom_image(data = calendar_image, aes(x = x, y = y, image = image), size = 0.18) + #Add calendar image in the coordinates set before 

  labs( #Clear labels and titles for the graph
    title = "Leap Day Birth Counts Through History", 
    x = "Year of Birth",
    y = "Number of People Born",
    caption = "Data Source: Leap Day TidyTuesday GitHub" #Where to find the data
  ) +
  
  theme_minimal() +
  theme( #
    panel.grid.major = element_line(color = "grey90"), #Makes grid lines light grey
                #easy to see
    panel.grid.minor = element_blank(), #No minor grid lines 
                #Makes it cleaner
    axis.text.x = element_text(angle = 45, hjust = 1), #Changing x-axis text
                #Helps fit more labels
    panel.border = element_rect(color = "grey70", fill = NA) #Adds a border around the plot
  ) +
  
  scale_x_continuous( #Scale for the x axis
    breaks = sort(c(seq(min(births_count$Birth_year), 2000, by = 50), #Adds earliest birth year then up by 50 years each line
                    1984, #Adds 1984 year cause that year had the maximum births
                    2004)), #Adds 2004 which is the last year with data
    limits = c(min(births_count$Birth_year), 2004) #Limits of the x-axis
  ) +
  
  scale_y_continuous(limits = c(1, 9), breaks = 1:9) #Scale for the y axis 
                # Has enough space for the image at the top

ggsave(here("Good_plot_Bad_plot","Output","Goodplot.pdf"), plot = good_plot, width = 8, height = 6) #save plot

print(good_plot)
```

[Summary of good plot:]{.underline}

The plot shows the number of Leap Day births over time. The maximum value of births is highlighted with a red dot and labeled, making it easy to identify that the peak occurred in 1984 with 8 births.
