---
title: "Week_07_OnlineVideo"
author: "Leah Barkai"
date: today
format: html
editor: visual
---

---
extra stuff I could use:
knitr: #this is to save figures automatically 
  opts_chunk:
    fig.path: "../Output/" ## added the ../ because your qmd folder is in your scripts folder so we need to go up one directory and then into your output folder
    message: false #gets rid of messages from code
    warning: false #gets rid of warnings from code
---

## Introduction

This document is for all the stuff I did while watching the online video lecture for week 7.

# [DepthData has:]{.underline}

-   [Date (YYYY-MM-DD HH:MM:SS)]{.underline}

-   [Atmospheric Pressure (dbar)]{.underline}

-   [Depth (meters)]{.underline}

# Load Libraries

These are the libraries I needed for this video.

for this homework assignment.

```{r}
library(ggmap)
library(tidyverse) #loads a bunch of data-science packages
library(here) #helps to call in data
library(readr) #to load in the data
library(ggspatial)



----
library(tidyverse) #loads a bunch of data-science packages
library(here) #helps to call in data
library(knitr) #lets you run code chunks in Quarto and R Markdown
library(kableExtra) #For kbl() and kable_styling()
library(ggplot2)


```

# Read in data

This data is from Maunalua Chemistry data.

```{r}
ChemData<-read_csv(here("Week_07","Data","chemicaldata_maunalua.csv")) #adding the data in
```

# Look at the data

This was to get an idea of the format of the data.

```{r}
glimpse(Chemd) #look at the structure, types and first few rows of the data table
```

# Data for Data Analysis

```{r}
Oahu<-get_map("Oahu")
ggmap(Oahu)
```

```{r}
#Make a data frame of lon and lat coordinates
WP<-data.frame(lon = -157.7621, lat = 21.27427) # coordinates for Wailupe
# Get base layer
Map1<-get_map(WP)
# plot it
ggmap(Map1)

Map1<-get_map(WP,zoom = 17)
ggmap(Map1)

Map1<-get_map(WP,zoom = 17, maptype = "satellite")
ggmap(Map1)

Map1<-get_map(WP,zoom = 17, maptype = "stamen_watercolor", source = "stadia")
ggmap(Map1)

Map1<-get_map(WP,zoom = 17, maptype = "satellite") 
ggmap(Map1)+
  geom_point(data = ChemData,
             aes(x = Long, y = Lat, color = Salinity),
             size = 4) +
  scale_color_viridis_c()

ggmap(Map1)+
  geom_point(data = ChemData, 
             aes(x = Long, y = Lat, color = Salinity), 
             size = 4) + 
  scale_color_viridis_c()+
  annotation_scale( bar_cols = c("yellow", "white"),
                    location = "bl")+ # put the bar on the bottom left and make the colors yellow and white
  annotation_north_arrow(location = "tl")+ # add a north arrow
  coord_sf(crs = 4326) # for the scale bar to work it needs to be in this coordinate system - this is a typical coordinate reference system for a GPS (WGS84)

geocode("the white house")

install.packages("emojifont")
library(emojifont)
search_emoji('smile')
## [1] "smiley"      "smile"       "sweat_smile" "smiley_cat"  "smile_cat"

ggplot() + 
  geom_emoji('smile_cat', 
             x=1:5, y=1:5, 
             size=10)



```
