---
title: "My quarto part 2"
author: "Leah Barkai"
date: today
format: 
  html:
    df-print: kable
knitr:
  opts_chunk:
    fig.path: "../Output/" ## added the ../ because your qmd folder is in your scripts folder so we need to go up one directory and then into your output folder
---
### Introduction 
Today I am going to plot some penguin data and learn how to work with figures. 
```{r}
#|message: false
#|warning: false

library(kableExtra)
library(tidyverse)
library(here)
library(palmerpenguins)
library(knitr)

```
## Look at the data 
```{r}
glimpse(penguins)
```
## Make a plot 
This is looking at the relationship between **bill length** and **bill depth**
```{r}
#| warning: false 
#| out-width: "70%"
#| fig-align: "center"
#| label: fig-penguin
#| fig-cap: "This is a figure showing bill length by bill depth"

penguins %>% 
  ggplot(aes(x=bill_length_mm, y=bill_depth_mm, color = species))+
  geom_point()
```
When you look at @fig-penguin we can see a positive relationship between bill length and bill depth

## Table

| Time  | Session | Topic    |
|:------|:-------:|---------:|
| _left_ | _center_ | _right_ |
| 01:00  | 1       | Anatomy |
| 01:50  |         | *Break* |
| 02:00  | 2       | Tables  |
| 02:45  |         | *Break* |

#Average bill length by species 
```{r}
#| warning: false
#| message: false
#| label: tbl-penguin
#| tbl-cap: This is my penguin table

penguins %>%
  group_by(species) %>%
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() %>%# make it a kable table
  kable_classic() %>% # add a theme that adds different colors 
  row_spec(2, bold = TRUE, color = "white", background = "red") %>% # highlight row 2 to have things stand out 
  kable_styling(full_width = FALSE) # don't make it so wide

```

Table @tbl-penguin has info on the mean bill length by species